---
title: "YET Zero Eye Tracking Data Analysis"
author: "M Schmettow"
date: "05/10/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.height = 8)
```

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(printr)

require(readxl)
require(jpeg)
require(ggimg)
```

## Input and preparation

Yet Zero produces output in CSV format, containing meta data and measured coordinates. 
Per session, one CSV file is created.
During preparation we read in all CSV files, combine them into a single data frame, and add external meta data on stimuli and areas of interest (AOI).

For this process, the following standard files and folders are used:

1. A folder "Data/" containing all CSV files produced by Yet Zero
1. A folder "Stimuli/" containing all stimulus images
1. A file "Stimuli/Stimuli.csv" containing meta data on the stimuli
1. A file "Stimuli/AOI.csv" containing definitions of areas of interest (AOI) on the stimuli

We begin with reading in in th stimulus data and how this can be put into static ggplot figures. 
This is particularly useful for checking the AOI definitions visually. 

### Reading Stimulus meta data

The following function reads in the stimulus meta data from the specified CSV file, and adds the image paths and image data.
The `col_types = col(...)` argument specifies the expected data types of each column in the CSV file. Always remember that CSV data is of type Character, including the numbers. `read_csv` often guesses the data type correctly, but there exist fringe cases, where this fails.


```{r}
col_types_Stim <- cols(
  File = col_character(),
  width = col_double(),
  height = col_double()
)

read_stim <- function(
  file = "Stimuli/Stimuli.csv", 
  stim_dir = "Stimuli/"
) {
  
  stimuli <- read_csv(file, col_types = col_types_Stim) %>%
    
    # Clean filenames 
    mutate(
      File = str_replace(File, ".jpg", ".png"), # remove hidden spaces 
      Path = str_c(stim_dir, File),
      Stim = File
    ) 
      
  
  return(stimuli)
}

Stimuli <- read_stim()

```

Here we use `read_stim` to create the Stimulus table. We do a little recoding, if needed. The file name acts as an *identifier* for the stimuli.

```{r}
Stimuli <- read_stim() %>% 
  mutate(Stim = File)
```

### Making a stimulus grid

A useful way to visualize eye tracking data is to use a grid of the stimuli (like a thumbnail sheet) and overlay the eye tracking data on top of it. This can take different forms, but the underlying grid is always the same. Here we create a re-usable ggplot object G_0 containing a grid of pictures.

```{r plot_stimuli, fig.height = 8, fig.width = 8}

#important for waldo_12 since it is a png but named as a jpg
Stimuli <- Stimuli %>%
  mutate(
    File = str_replace(File, ".jpg", ".png"),
    Path = str_replace(Path, ".jpg", ".png")
  )

library(ggimage)

G_0 <- 
  Stimuli %>% 
  ggplot(aes(x = width/2, y = height/2)) +
  facet_wrap(~Stim) +
  ggimage::geom_image(aes(image = Path), size = 1)

G_0

```

### Reading csv

The following code contains two functions. The first function `read_yz_csv` reads in a single CSV file, removes duplicate rows (which can occur when the eye tracker loses track of the eye).
The second function `read_yz_files` takes a *list* of file names and reads them, combines them into a single data frame, and adds stimulus meta data from the stimulus table created above. 

** Notes **

The `map_df()` function is a so called *iterator*. Iterators are called *super functions*, where one of the function arguments is a function itself. In R iterators are the default way to do an operation repeatedly over a data set.
Now you know why tidy R never uses for loops, like in Python. And you know one reason why R is called a *functional programming language*!


There is another famous piece of data science at work here. 
Think about the problem of adding the Stimulus meta data to the eye tracking data. 
Try to picture how you would do it manually: 

1. Identify matching Stimulus ID in both tables
2. Copy the meta data from the Stimulus table to the eye tracking data table
3. Repeat for all rows in the eye tracking data table.

This operation is called a *join* in Data Science. 
A precondition for joins is that the table have matching identifiers to create the relationship.
Here this is the `Stim` column.


```{r read_data}
csv_files <- dir(
  path = "Data/",
  pattern = "*.csv",
  full.names = TRUE
)

read_yz_csv <- function(file){
  read_csv(file, 
           col_types = cols(
             Part = col_character(), 
             Stim = col_character(), 
             time = col_double(), 
             xF = col_double(), 
             yF = col_double(), 
             xF_pro = col_double(), 
             yF_pro = col_double()
           )) %>% 
    
    # clean Stim names (VERY IMPORTANT)
    mutate(
      Stim = str_trim(Stim)
    ) %>%
    
    # remove duplicates (based on gaze position)
    mutate(is_duplicate = xF == lag(xF) & yF == lag(yF)) %>% 
    filter(!is_duplicate) %>% 
    mutate(File = file) %>% 
    
    # rename to match rest of pipeline
      rename(
      x = xF,
      y = yF,
      x_pro = xF_pro,
      y_pro = yF_pro
    ) %>% 
    select(Part, Stim, time, x, y, x_pro, y_pro)
}

read_yz_data <- function(files, stim_tab){
  Data <- files %>% 
    map_df(read_yz_csv) %>% 
    
    mutate(Obs  = row_number()) %>%
    
    mutate(Part = as.factor(as.integer(Part) - min(as.integer(Part)) + 1)) %>%
    
    group_by(Part) %>%
    mutate(time = time - min(time)) %>%
    ungroup() %>%
    
    left_join(stim_tab, by = "Stim") %>% 
    
     #check join success (debugging helper)
    mutate(join_ok = !is.na(width)) %>%
    
    mutate(
      y = height - y,   
      y_pro = 1 - y_pro
    )
  
  return(Data)
}
```

Before we can use this function, we have to create a list of all Csv files we want to process.
The shown default is to read all csv files from directory `Data`. 

```{r reading_csv}
csv_files <- dir(
  path = "Data/",
  pattern = "*.csv",
  full.names = TRUE
)

# read and combine all data
D_0 <- read_yz_data(csv_files, Stimuli)

# keep only valid gaze points inside screen
D_in <- D_0 %>% 
  filter(
    !is.na(x), !is.na(y),          # remove missing gaze points
    between(x, 0, width),
    between(y, 0, height)
  )

head(D_in)

```

# Last Session

Before continue with AOI coding, we take a look at the data of the last participant only. This is useful to check right after an experiment.

Here we also introduce some basic visualizations of raw data, or overlays to the grid of Stimuli. The following function extracts the data of the last participant only.


```{r}
get_last_part <- function(data){
  
  last_part_id <- max(as.numeric(as.character(data$Part)), na.rm = TRUE)
  
  last_part <- data %>%
    filter(as.numeric(as.character(Part)) == last_part_id)
  
  return(last_part)
}
```


Here we extend the stimulus grid G_0 with points of measurement. See how the `geom_point` function takes a `data` argument, which allows us to specify a different data frame than the one used for the base ggplot object G_0 (which is the table of stimuli).

Note that the individual figures are being scaled by the eye tracking measures.

```{r heatmap_last, fig.width=12, fig.height=10}
G_0 +
  geom_point(
    data = get_last_part(D_in),   # use cleaned data
    aes(x = x, y = y),
    size = 1.5,
    col = "red",
    alpha = 0.3
  )
```
 The following variation shows the path of the last participant, by connecting the points with lines.

```{r pathmap_last, fig.width=12, fig.height=10}
G_0 +
  geom_point(
    data = get_last_part(D_in),
    aes(x = x, y = y),
    size = 0.3,
    alpha = 0.4
  ) +
  geom_line(
    data = get_last_part(D_in),
    aes(x = x, y = y, group = Part),
    col = "red",
    alpha = 0.7
  )
```

Making an orderly scan path based on time for waldo_05 as an example

```{r}
library(tidyverse)
library(ggimage)

# Select participant and stimulus
stim_name <- "waldo_05.png"
part_id <- max(as.numeric(as.character(D_in$Part)))

scan_data <- D_in %>%
  filter(Part == part_id, Stim == stim_name) %>%
  arrange(time) %>%
  mutate(order = row_number())

# Get stimulus dimensions
stim_info <- Stimuli %>%
  filter(Stim == stim_name)

# Create the scan path plot
ggplot() +
  # Display the stimulus image
  ggimage::geom_image(
    data = stim_info,
    aes(x = width / 2, y = height / 2, image = Path),
    size = 1
  ) +
  
  # Draw scan path lines
  geom_path(
    data = scan_data,
    aes(x = x, y = y),
    color = "red",
    linewidth = 1.2,
    alpha = 0.8
  ) +
  
  # Draw gaze points colored by fixation order
  geom_point(
    data = scan_data,
    aes(x = x, y = y, color = order),
    size = 3
  ) +
  
  # Optional: label the first few fixations
  geom_text(
    data = scan_data %>% filter(order <= 10),
    aes(x = x, y = y, label = order),
    color = "black",
    size = 3,
    vjust = -1
  ) +
  
  # Use a perceptually uniform color scale
  scale_color_viridis_c(option = "plasma") +
  
  # Constrain the plot to the stimulus dimensions
  coord_fixed(
    xlim = c(0, stim_info$width),
    ylim = c(0, stim_info$height),
    expand = FALSE,
    clip = "on"
  ) +
  
  labs(
    title = paste("Scan Path for", stim_name),
    subtitle = "Gaze points colored by fixation order",
    x = "X position (pixels)",
    y = "Y position (pixels)",
    color = "Fixation\nOrder"
  ) +
  
  theme_minimal() +
  theme(
    plot.margin = margin(0, 0, 0, 0)
  )

```



Here, we use a code to color by fixation order. By default, ggplot uses a continuous color scale,meaning that low values are represented by yellow and high values are represented by dark blue. The plot showcases the start of the search (dark blue), the middle of the search (green), and the end of the search (yellow). 


```{r fixationorder_last, fig.width=12, fig.height=10}
# Determine the actual presentation order for each participant
stim_order <- D_in %>%
  group_by(Part, Stim) %>%
  summarise(first_time = min(time), .groups = "drop") %>%
  arrange(Part, first_time) %>%
  group_by(Part) %>%
  mutate(trial_order = row_number()) %>%
  ungroup()

# Join the presentation order with the stimulus metadata
Stimuli_ordered <- Stimuli %>%
  left_join(stim_order %>% select(Part, Stim, trial_order),
            by = "Stim")

# Get presentation order for the last participant
last_part <- max(as.numeric(as.character(D_in$Part)))

stim_levels <- stim_order %>%
  filter(Part == last_part) %>%
  arrange(trial_order) %>%
  pull(Stim)

# Apply ordered factor to the data used in the plot
plot_data <- get_last_part(D_in) %>%
  mutate(Stim = factor(Stim, levels = stim_levels)) %>%
  arrange(Stim, time) %>%
  group_by(Stim) %>%
  mutate(order = row_number()) %>%
  ungroup()

Stimuli_plot <- Stimuli %>%
  mutate(Stim = factor(Stim, levels = stim_levels))

# Create the ordered fixation plot
G_ordered <- 
  Stimuli_plot %>%
  ggplot(aes(x = width / 2, y = height / 2)) +
  facet_wrap(~Stim) +
  ggimage::geom_image(aes(image = Path), size = 1)

G_ordered +
  geom_point(
    data = plot_data,
    aes(x = x, y = y, col = order),
    size = 1.5,
    alpha = 0.6
  ) +
  scale_color_viridis_c(name = "Fixation Order") +
  labs(title = "Fixation Order in Actual Stimulus Presentation Sequence")

```

Additional plot to show how search develops over time, each color represents 1/3 of the search time, showing beginning, middle, and end of the search represented by colors 

```{r search_behavior, fig.width=12, fig.height=10}
D_in %>%
  group_by(Part, Stim) %>%
  arrange(time) %>%
  mutate(
    phase = case_when(
      row_number() < n()/3 ~ "early",
      row_number() < 2*n()/3 ~ "middle",
      TRUE ~ "late"
    )
  ) %>%
  ggplot(aes(x = x, y = y, col = phase)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~Stim)
```

msking the plot again but specifically for waldo_05

```{r}
stim_select <- "waldo_05.png"

# base plot with only one stimulus
G_single <- 
  Stimuli %>% 
  filter(Stim == stim_select) %>% 
  ggplot(aes(x = width/2, y = height/2)) +
  ggimage::geom_image(aes(image = Path), size = 1)

# add your phase-colored gaze points
G_single +
  geom_point(
    data = D_in %>%
      group_by(Part, Stim) %>%
      arrange(time) %>%
      mutate(
        phase = case_when(
          row_number() < n()/3 ~ "early",
          row_number() < 2*n()/3 ~ "middle",
          TRUE ~ "late"
        )
      ) %>%
      ungroup() %>%
      filter(Stim == stim_select),
    
    aes(x = x, y = y, col = phase),
    size = 7.5,      # make points visible
    alpha = 2.0
  )
```

The following code creates a density plot, it groups gaze points into small squares (bins) across the image instead of showing every single gaze point. It shows exactly how many gaze points fall into each area without creating overlap. light squares mean few or no gaze points, dark squares mean many gaze points. 

```{r densityplot_last, fig.width=12, fig.height=10}
#density plot
G_0 +
  geom_bin2d(
    data = D_in,
    aes(x = x, y = y),
    bins = 40
  ) +
  facet_wrap(~Stim)
```

## Deriving measures

Often, it is relevant to know where someone is looking at, but not always. Here we derive a simple measure, the total distance their eyes travelled during stimulus presentation.

```{r}
add_travel <- function(data) {
  mutate(data,
         travel = sqrt((x - lag(x))^2 + (y - lag(y))^2))
}

D_1 <- D_in %>%   # use cleaned data
  mutate(Sequence = paste(Part, Stim, sep = "_")) %>% 
  
  group_by(Sequence) %>% 
  arrange(time) %>%   
  
  add_travel() %>% 
  
  ungroup() %>% 
  
  select(Obs, Part, Stim, Sequence, time, x, y, travel)
```

```{r plot_travel}

D_1 %>% 
  group_by(Stim) %>% 
  summarise(mean_travel = mean(travel, na.rm = TRUE)) %>% 
  
  ggplot(aes(x = as.numeric(as.factor(Stim)), y = mean_travel)) +
  geom_line() +
  labs(x = "Stimulus order", y = "Mean eye movement distance")

```

#Performance measures without AOIs

Reaction time   

```{r}
D_perf <- D_in %>%
  group_by(Part, Stim) %>%
  
  summarise(
    rt = max(time) - min(time),
    start_time = min(time),   # capture when trial started
    .groups = "drop"
  ) %>%
  
  mutate(
    rt = pmin(rt, 45),
    accuracy = if_else(rt < 45, 1, 0)
  ) %>%
  
  arrange(Part, start_time) %>%   # TRUE order!
  
  group_by(Part) %>%
  mutate(trial = row_number()) %>%
  ungroup()

```


```{r}

ggplot(D_perf, aes(x = trial)) +
  
  # background bars = accuracy
  geom_col(aes(y = accuracy * 45),
           fill = "grey80",
           alpha = 0.5) +
  
  # individual participant lines
  geom_line(aes(y = rt, group = Part), color = "black", alpha = 0.4) +
  geom_point(aes(y = rt), color = "black", alpha = 0.6) +
  
  # highlight DNF trials
  geom_point(
    data = D_perf %>% filter(rt == 45),
    aes(y = rt),
    color = "red",
    size = 3
  ) +
  
  # cutoff line
  geom_hline(yintercept = 45, linetype = "dashed", color = "red") +
  
  # TREND LINE (mean trend across all participants)
  geom_smooth(
    aes(y = rt),
    method = "loess",   # smooth curve (nice for behavior)
    se = FALSE,
    color = "blue",
    size = 1.2
  ) +
  
  labs(
    title = "Performance Over Trials",
    subtitle = "Reaction Time with Learning/Fatigue Trend",
    x = "Trial",
    y = "Reaction Time (seconds)"
  )
```

Reaction time combined with Accuracy 

```{r}
library(tidyverse)

# Read the accuracy data
accuracy <- read_csv("accuracy.csv")

# Rename columns for easier handling
accuracy <- accuracy %>%
  rename(
    trial = `trial-index`,
    Stim = stimulus,
    Accuracy = clicked_in_waldo,
    rt = time
  )

# If the dataset contains only one participant, create a participant ID
accuracy <- accuracy %>%
  mutate(Part = 1)

# Ensure the data is ordered according to the actual presentation sequence
accuracy <- accuracy %>%
  arrange(Part, trial)
```

```{r}
ggplot(accuracy, aes(x = trial)) +
  
  # Background bars representing accuracy (scaled to RT axis)
  geom_col(
    aes(y = if_else(Accuracy == 1, 45, 0)),
    fill = "grey80",
    alpha = 0.5
  ) +
  
  # Reaction time line
  geom_line(aes(y = rt, group = Part), color = "black", linewidth = 1) +
  geom_point(aes(y = rt), color = "black", size = 2) +
  
  # Highlight DNF trials (timeouts)
  geom_point(
    data = accuracy %>% filter(Accuracy == -1),
    aes(y = rt),
    color = "red",
    size = 3
  ) +
  
  # Highlight incorrect clicks
  geom_point(
    data = accuracy %>% filter(Accuracy == 0),
    aes(y = rt),
    color = "orange",
    size = 3
  ) +
  
  # 45-second cutoff line
  geom_hline(yintercept = 45, linetype = "dashed", color = "red") +
  
  # Trend line (learning/fatigue)
  geom_smooth(
    aes(y = rt),
    method = "loess",
    se = TRUE,
    color = "blue",
    linewidth = 1.2
  ) +
  
  labs(
    title = "Performance Across Trials",
    subtitle = "Reaction Time and Accuracy in the Waldo Task",
    x = "Trial (Presentation Order)",
    y = "Reaction Time (seconds)"
  ) +
  theme_minimal()
```

```{r}
accuracy_summary <- accuracy %>%
  group_by(trial, Accuracy) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(trial) %>%
  mutate(prop = n / sum(n))

ggplot(accuracy_summary, aes(x = trial, y = prop, fill = factor(Accuracy))) +
  geom_area(position = "stack", alpha = 0.8) +
  scale_fill_manual(
    values = c("-1" = "red", "0" = "orange", "1" = "darkgreen"),
    labels = c("-1" = "DNF", "0" = "Incorrect", "1" = "Correct"),
    name = "Outcome"
  ) +
  labs(
    title = "Proportion of Trial Outcomes Over Time",
    x = "Trial (Presentation Order)",
    y = "Proportion"
  ) +
  theme_minimal()
```

```{r}
ggplot(accuracy, aes(x = reorder(Stim, rt, median), y = rt, fill = factor(Accuracy))) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_manual(
    values = c("-1" = "red", "0" = "orange", "1" = "darkgreen"),
    labels = c("-1" = "DNF", "0" = "Incorrect", "1" = "Correct"),
    name = "Outcome"
  ) +
  coord_flip() +
  labs(
    title = "Reaction Time by Stimulus",
    x = "Stimulus",
    y = "Reaction Time (seconds)"
  ) +
  theme_minimal()
```

THE AOI PLOTS WERE NOT USED IN R DUE TO NAMING/FORMATTING ISSUES
-> the analysis continued in thonny

# Areas of interest

The main use case for eye tracking is to measure how often or how long, or in which sequence, people look at certain areas of interest (AOI) on the stimulus. What these areas are differs from case to case, here it is the main areas of faces: eyes, mouth and nose.

AOIs are defined as rectangles on the stimulus images. The following function reads in the AOI definitions from a CSV file, and merges them with the stimulus meta data to create absolute coordinates for each AOI on each stimulus.


## Reading AOI

The function defined here makes againn use of a join operation, 
which merges the stimulus meta data with the AOI rectangle definitions. 
Different to the previous join, the linking iodentifier is not the Stimulus ID, but the face. This is a difference, as in this experiment, the same face appeared in two conditions regarding the sclera color. This is also the reason why the join is many-to-many.

```{r}
read_yz_aoi <- function(file = "Stimuli/AOI.csv",
                        col_types = cols(
                          Stimulus = col_character(),
                          AOI = col_character(),
                          x = col_double(),
                          y = col_double(),
                          w = col_double(),
                          h = col_double()
                        ),
                        stim_tab) {
  
  AOI <- read_csv(file, col_types = col_types) %>% 
    
    
    
    # rename columns
    rename(
      Stim = Stimulus,  
      x_aoi = x,
      y_aoi = y,
      w_aoi = w,
      h_aoi = h
    ) %>% 
    
    # normalize stimulus names (remove extensions everywhere)
    mutate(
      Stim = str_remove(Stim, "\\.(jpg|png)$")
    )
  
  Stim_clean <- stim_tab %>% 
    mutate(
      Stim = str_remove(Stim, "\\.(jpg|png)$")
    )
  
  # join with stimulus metadata
  AOI <- AOI %>%
    left_join(Stim_clean, by = "Stim") %>% 
    
    # compute AOI boundaries (important for correct plotting!)
    mutate(
      join_ok = !is.na(width),
      
      xmin = x_aoi,
      xmax = x_aoi + w_aoi,
      
      ymax = height - y_aoi,
      ymin = ymax - h_aoi
    ) %>% 
    
    arrange(Stim, AOI)
  
# debugging 
  
if(any(!AOI$join_ok)) {
    warning("⚠️ Some AOIs did not match stimuli!")
    print(unique(AOI$Stim[!AOI$join_ok]))
  }
  
  return(AOI)
}
```


```{r reading_aoi}
AOI <- read_yz_aoi(stim_tab = Stimuli)

unique(AOI$Stim)
unique(Stimuli$Stim)

```

## AOI preview

Like we did before with the stimulus images, we can now create a ggplot object that shows the AOI rectangles on top of the stimulus images. This is very useful for checking if the AOI definitions are correct.

```{r}

G_1 <- 
  AOI %>% 
  ggplot(aes(x = width/2, y = height/2)) +
  
  facet_wrap(~Stim) +
  
  ggimage::geom_image(aes(image = Path), size = 1) +
  
  geom_rect(
    aes(xmin = xmin, ymin = ymin, 
        xmax = xmax, ymax = ymax,
        fill = AOI),
    alpha = .2,
    inherit.aes = FALSE
  )

G_1
```

#### AOI Classification

We have the AOI definitions, now we can classify each eye tracking observation according to the AOI it falls into. For this purpose, we do a many-to-many join between the eye tracking data and the AOI definitions.

But this is just an intermediate step. Next, we filter only those rows, where the measure falls into the AOI. As long as the AOIs are non-overlapping, there can only be one (or zero) match. Then we check for each observation if it falls into the rectangle defined by the AOI. If so, we keep that observation-AOI combination. Finally, we right-join this result with the original eye tracking data to keep all observations, and assign "Outside" to those observations that did not fall into any AOI.


```{r}
D_2 <- 
  D_in %>% 
  
  #join by stimulus
  left_join(AOI, by = "Stim", relationship = "many-to-many") %>% 
  
  # check if gaze point is inside AOI
  mutate(is_in = x > xmin & x < xmax & y > ymin & y < ymax) %>% 
  
  filter(is_in) %>% 
  
  select(Obs, AOI) %>% 
  
  # keep all observations
  right_join(D_1, by = "Obs") %>% 
  
  # keep screen boundaries
  left_join(D_in %>% distinct(Obs, width, height), by = "Obs") %>% 
  
  filter(between(x, 0, width), between(y, 0, height)) %>% 
  
  # label outside points
  mutate(AOI = if_else(is.na(AOI), "Outside", AOI)) %>% 
  
  arrange(Part, Stim, time)
```


```{r fig.width=12, fig.height=10}
# AOI counts per stimulus 

D_2 %>% 
  group_by(Stim, AOI) %>% 
  summarize(n = n(), .groups = "drop") %>% 
  
  ggplot(aes(x = AOI, y = n, fill = AOI)) +
  geom_col() +
  facet_wrap(~Stim)

```

```{r plot_aoi_counts, fig.width=12, fig.height=10}
G_0 +
  geom_count(
    aes(x = x, y = y, col = AOI),
    alpha = .5,
    inherit.aes  = FALSE,
    data = D_2
  )
```
### Performance measures with AOIs

possible codes to measure performance in terms of RT and accuracy based on the AOIs (waldo AOIs have to be added to the csv)

```{r}
#reaction time finding waldo

RT <- D_2 %>% 
filter(AOI == "Waldo") %>% 
group_by(Part, Stim) %>% 
summarize(
   RT = min(time),
    .groups = "drop"
)
```

```{r}
#accuracy 

RT <- D_2 %>% 
filter(AOI == "Waldo") %>% 
group_by(Part, Stim) %>% 
summarize(
RT = min(time),
  .groups = "drop"
  )
```

### Refixations (fatigue)

```{r}
D_2 %>% 
  group_by(Part, Stim, AOI) %>% 
  summarize(n_fix = n()) %>% 
  filter(AOI == "Waldo")
```

### Measuring visits

A *visit* is a closed sequence of eye positions in the same region. The following code uses a combined criterion for setting a new visits:

-   the position falls into a different AOI

-   OR: the travel traveled from the previous position exceeds a certain threshold

```{r}
travel_threshold <- 50

D_3 <-
  D_2 %>%
  group_by(Part, Stim) %>%
  arrange(time) %>% 
  
  mutate(
    new_visit = AOI != lag(AOI) | travel > travel_threshold,
    
    # first row should always start a visit
    new_visit = if_else(is.na(new_visit), TRUE, new_visit),
    
    visit = cumsum(new_visit),
    
    duration = lead(time) - time
  ) %>%
  
  ungroup()

```

#### Plotting visit paths and duration

```{r }

Visit_summary <- D_3 %>%
  group_by(Part, Stim, visit, AOI) %>%
  summarize(
    x = mean(x),
    y = mean(y),
    duration = sum(duration, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r fig.width=8, fig.height = 8}
# plot 1 visit duration (fatigue & attention)
G_3 <- 
  G_0 +
  geom_point(
    data = Visit_summary %>% filter(AOI != "Outside"),
    aes(x = x, y = y, size = duration, col = AOI),
    alpha = 0.7,
    inherit.aes = FALSE
  )

G_3
```

```{r fig.width=8, fig.height = 8}
# plot 2 search paths 
library(grid)  # for arrows

G_4 <- 
  G_0 +
  geom_path(
    data = Visit_summary %>% filter(AOI != "Outside"),
    aes(x = x, y = y, group = interaction(Part, Stim), col = AOI),
    arrow = arrow(length = unit(0.2, "cm")),
    alpha = 0.8,
    inherit.aes = FALSE
  ) +
  geom_text(
    data = Visit_summary %>% filter(AOI != "Outside"),
    aes(x = x, y = y, label = visit),
    size = 3,
    inherit.aes = FALSE
  )

G_4

```

## Participant-level analysis

### Frequencies and durations

```{r}

D_4 <-
  D_3 %>%
  group_by(Part, face, AOI, Sunglasses) %>%  # <--
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>%
  ungroup() #%>% 
#  mutate(congruent = (Sclera == Skull))


D_4

```

```{r}

#G_6 <-
#
#  D_4 %>%
#  ggplot(aes(x = congruent, y = n_visits, fill = AOI)) +
#  facet_wrap(~Part) +
#  geom_col()
#
#G_6
#
```


```{r}

#G_7 <-
#  D_4 %>%
#  ggplot(aes(x = AOI, y = total_dur, fill = congruent)) +
#  facet_wrap(~Part) +
#  geom_col()
#
#G_7

```


